---
title: "Homework 3 - P8105"
author: "Sergio Ozoria Ramírez (smo2147)"
date: "2025-10-02"
output: github_document
---

# Problem 1

#### Setup & Loading Instacart Dataset

```{r loading packages + setup, message = FALSE}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 1,
  out.width = "100%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#### Description of Instacart Dataset

The instacart dataset describes online grocery shopping behaviors from local stores across New York City from 2017. The dataset contains **`r nrow(instacart)`** observations and **`r ncol(instacart)`** variables in total, including the name of the product (i.e., `product name`), the aisle (i.e., `aisle`), and the department (i.e., `department`). The dataset also includes a mix of numeric and categorical variables. The average order hour of the day among shopping users is approximately **`r instacart |> pull(order_hour_of_day) |> mean(na.rm = TRUE) |> round()`**, whereas the average number of days since last order is **`r instacart |> pull(days_since_prior_order) |> mean(na.rm = TRUE) |> round()`**. Altogether, this means users mostly grocery shop early afternoon with users typically reordering every 17 days.

There are `r instacart |> pull(aisle) |>  n_distinct()` unique aisles in this dataset, with **`r instacart |> count(aisle, sort = TRUE) |> slice_head(n = 10) |> pull(aisle)`** being the top 10 aisles where items are mostly ordered from.

#### Figure.1. Bar Chart of Number of Items Ordered by Aisle

```{r}
instacart |> 
  count(aisle, sort = TRUE) |> 
  filter(n > 10000) |> 
  ggplot(aes(x = reorder(aisle, n), y = n)) +
  labs(
    title = "Aisles with More than 10,000 Items Ordered",
    x = "Aisle Categories",
    y = "Number of Aisles Ordered",
    caption = "Data: Instacart Online Grocery Shopping (2017)" 
  ) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(
    breaks = c(0, 25000, 50000, 75000, 
               100000, 125000, 150000))
```

#### Table.1. Most Popular Items in Aisles `Baking Ingredients`, `Dog Food Care`, and `Packaged Vegetables Fruits`

```{r}
instacart |> 
  filter(aisle %in% c(
    "packaged vegetables fruits",
    "baking ingredients", 
    "dog food care")) |>
  count(aisle, product_name, name = "total_ordered", sort = TRUE) |> 
  group_by(aisle) |> 
  slice_max(total_ordered, n = 3) |> 
   rename(
    Aisle = aisle,
    Product = product_name,
    `# of Items Ordered` = total_ordered
  ) |> 
  knitr::kable()
```

```{r}
instacart |> 
  filter(product_name %in% c(
    "Pink Lady Apples",
    "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = round(mean(order_hour_of_day, na.rm = TRUE))) |> 
  mutate(order_dow = recode(order_dow,
   `0` = "Sunday",
   `1` = "Monday",
   `2` = "Tuesday",
   `3` = "Wednesday",
   `4` = "Thursday",
   `5` = "Friday",
   `6` = "Saturday",
  )) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable()
```

